---
title: "processROV"
author: "Kevin L. Stierhoff"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
# Install and load pacman (library management package)
if (!require("pacman")) install.packages("pacman")

# Install and load required packages from CRAN ---------------------------------
pacman::p_load(tidyverse,lubridate,here,RODBC,forecast,cowplot,exifr,sf,fs)

# Install and load required packages from Github -------------------------------
# surveyR
pacman::p_load_gh("kstierhoff/surveyR")

# Set default ggplot2 theme
theme_set(theme_bw())

# Create required directories
dir_create(here(c("Data","Figs","Output")))
```

# Load user settings

User-defined settings control various processing steps and outputs.

```{r user-settings}
copy.files        <- T # Copy files from root directories
process.files     <- T # Process files
process.files.all <- T # Process all files, not only new ones
process.photos    <- T # Process photos
```

# Load project settings

Load the project-specific settings. The appropriate file is determined by the name of the directory containing the Rstudio project.  
```{r project-settings}
# Get project name from directory
prj.name <- last(unlist(str_split(here(),"/")))

# Get all settings files
settings.files <- dir(here("Doc/settings"))

# Source survey settings file
prj.settings <- settings.files[str_detect(settings.files, paste0("settings_", prj.name, ".R"))]
source(here("Doc/settings", prj.settings))
```

# Get database values

Query the database tables to determine the next index for each table, which is the "seed" value for new indices.

```{r get-seeds}
# Query starting IDs for all database tables ####
channel <- odbcConnectAccess2007(db.dir)
nav.seed    <- as.numeric(sqlQuery(channel, "SELECT max(dbo_tbl_NAV.nav_id) AS nav_seed
                                   FROM dbo_tbl_NAV;")[1] + 1)
event.seed  <- as.numeric(sqlQuery(channel, "SELECT max(dbo_tbl_EVENTS.event_id) AS event_seed
                                   FROM dbo_tbl_EVENTS;")[1] + 1)
photo.seed  <- as.numeric(sqlQuery(channel, "SELECT max(dbo_tbl_PHOTO_INFO.photo_id) AS photo_seed
                                   FROM dbo_tbl_PHOTO_INFO;")[1] + 1)
ctd.seed    <- as.numeric(sqlQuery(channel, "SELECT max(dbo_tbl_CTD_CASTS.ctd_id) AS ctd_seed
                                   FROM dbo_tbl_CTD_CASTS;")[1] + 1)
close(channel)
```

# Get directories

```{r get-directories}
# Get a list of WinFrog project directories that contain NAV data
d <- dir_ls(dat.root, regexp = "\\d{2}-\\d{3}\\w")
# Get a list of directories to process
dir.list <- d[grep(dir.start, d):grep(dir.end, d)]
# Manually remove directories to not be processed
dir.list <- dir.list[!path_file(dir.list) %in% dir.rm]

if (process.photos) {
  # Get a list of directories that contain photos
  p <- dir_ls(photo.root, regexp = "\\d{2}-\\d{3}\\w")
  # Get a list of directories to process
  photo.list <- p[path_file(p) %in% path_file(dir.list)] 
}
```

# Copy files

Copy WinFrog navigation data and photos to a local directory for processing, leaving the original files in place.

```{r copy-files}
if (copy.files) {
  # Copy WinFrog files
  dir_copy(dir.list, here("Data/NAVDATA", path_file(dir.list)), overwrite = TRUE)

  if (process.photos) {
    # Copy photos
    dir_copy(photo.list, here("Data/PHOTOS", path_file(photo.list)), overwrite = TRUE)
  }
}
```

# Process nav files

Extract data from WinFrog navigation files for each transect. Optionally, to speed processing, only newly added files may be processed (`process.all` = `TRUE`).

```{r process-nav}
if (process.files) {
  # List local DAT directories
  dat.dirs <- dir_ls(here("Data/NAVDATA"))
  
  if (!process.files.all) {
    # Load already processed DAT file directories 
    if (file.exists(here("Output/processed_dat.rds"))) {
      processed.dat <- readRDS(here("Output/processed_dat.rds"))
    }
    
    # Load already processed nav data
    readRDS(here("Data/NAVDATA/DAT.rds"))
  }  
  
  # Process DAT files
  if (!process.files.all) {
    # Get unprocessed directories
    dat.dirs <- dat.dirs[!dat.dirs %in% processed.dat]
    
    if (file.exists(here("Data/NAVDATA/DAT.rds"))) {
      # Load already processed vessel NASC data
      DAT <- readRDS(here("Data/NAVDATA/DAT.rds"))
    } else {
      # Create new data frame
      DAT <- tibble()  
    }
  } else {
    # Create new data frame
    DAT   <- tibble()
    CTD   <- tibble()
    LOG   <- tibble()
    PHOTO <- tibble()
  }
  
  # Process remaining directories
  for (i in dat.dirs) {  
    # extract dive name from directory path
    dive.name <- 	str_extract(i, "\\d{2}-\\d{3}\\w")
    
    # List DAT files in directory
    dat.files <- dir_ls(i, regexp = "\\d{2}-\\d{3}\\w{1}(\\(\\d{3}-\\d{6}\\))?.DAT")
    
    # Create temporary dataframe for dive data
    DAT.tmp <- tibble()
    
    for (ii in dat.files) {
      # Read DAT file and append to previous
      dat     <- read_csv(ii, col_names = F)
      DAT.tmp <- bind_rows(DAT.tmp, dat)
    }
    
    if (nrow(DAT.tmp) > 0) {
      # Subset columns
      DAT.tmp <- select(DAT.tmp , c(1:47,65))
      
      # Add variable names to data frame
      names(DAT.tmp) <- c("oid","blank","date_time","lat_r","long_r","depth","northing_r","easting_r",
                          "blank","blank","blank","heading_r","cmg_r","speed_r","blank","blank","blank","blank",
                          "blank","blank","blank","blank","pitch","roll","temperature","conductivity",
                          "pressure","salinity","sound_vel","oxygen_conc","oxygen_sat","altitude",
                          "blank","blank","lat_s","long_s","blank","northing_s","easting_s","blank",
                          "blank","blank","heading_s","cmg_s","speed_s","blank","blank","blank")
      
      # Remove variables with no data
      DAT.tmp <- DAT.tmp[ , names(DAT.tmp) != "blank"]
      
      DAT.tmp <- DAT.tmp %>% 
        mutate(nav_id    = seq(nav.seed, nav.seed + n() - 1),
               object_id = oid + 1,
               dive_name = as.factor(dive.name),
               date_time = mdy_hms(date_time),
               lat_r     = winfrog2dd(lat_r),
               long_r    = winfrog2dd(long_r),
               lat_s     = winfrog2dd(lat_s),
               long_s    = winfrog2dd(long_s),
               time.int  = c(as.numeric(date_time[2:n()] - date_time[1:n() - 1]), 0),
               disp_s    = speed_s * 0.514444444 * time.int,
               disp_r    = speed_r * 0.514444444 * time.int,
               good      = 1)
      
      # Increment nav.seed
      nav.seed <- max(DAT.tmp$nav_id) + 1
      
      # Add DVL pitch/roll to position sensor to get the actual pitch/roll of the camera ####
      # Get directory of WinFrog RAW files to read
      raw.files <- dir_ls(i, regexp = "*.RAW")
      
      # Create tibble for raw data
      pr.data <- data.frame()
      
      for (k in raw.files) {
        pr.temp <- readLines(k)
        # Extract only PASHR sentences
        pr.temp <- pr.temp[grep("413-004-W,DVL [on HDHV]*", pr.temp)]
        # Split strings on commas
        pr.temp <- str_split(pr.temp, ",")
        # Convert list to data frame
        pr.temp <- do.call(rbind.data.frame, pr.temp)
        # Subset only columns with DVL data
        pr.temp <- pr.temp[ , 1:5]
        # Add names to data frame
        names(pr.temp) <- c("code","name","time","pitch","roll")
        # Convert numeric values from factor
        pr.temp <- mutate_at(pr.temp, c("time","pitch","roll"), ~as.numeric(as.character(.))) 
        # Convert time to datetime
        t.seed <- ymd_hms('1970-01-01 00:00:00')
        pr.temp <- mutate(pr.temp, date_time = t.seed + time)
        # Combine data frames
        pr.data <- bind_rows(pr.data, distinct(pr.temp))
      }
      
      # Match pitch and roll data to nav 
      pid <- numeric()
      pid.lag <- numeric()
      
      # Add DVL pitch to the pitch measured by the tilt tray position sensor
      if (nrow(pr.data) != nrow(DAT.tmp)) {
        # If RAW data are logged 'Continuously'
        for (jj in 1:nrow(DAT.tmp)) {
          # Calculate the time difference between jth video obs and each nav record
          time.diff	  <- abs(difftime(DAT.tmp$date_time[jj], pr.data$date_time, units = "secs"))
          # Get min index for merging later
          pid <- c(pid, which.min(time.diff))
          # Get time lag
          pid.lag <- c(pid.lag, as.numeric(time.diff[which.min(time.diff)]))
        }
        
        DAT.tmp$pitch <- DAT.tmp$pitch + round(pr.data$pitch[pid])
      } else { 
        # If RAW data are logged 'With Events'
        DAT.tmp$pitch <- DAT.tmp$pitch + round(pr.data$pitch)
      }
      
      # Write DVL pitch/roll data to CSV
      write_csv(pr.data, file.path(i, paste(dive.name, "PitchRollData.txt", sep = "_")),
                na = "-999")
      
      # Write DVL pitch/roll data to CSV
      write_csv(pr.data, here("Output", paste(dive.name, "PitchRollData.txt", sep = "_")),
                na = "-999")
      
      # Process CTD data
      if (!ctd.on) {  
        # If CTD not present, make data -999
        DAT.tmp <- DAT.tmp %>% 
          mutate(
            sal           = -999,
            conductivity  = -999,
            oxygen_conc   = -999,
            oxygen_sat    = -999,
            depth_p_r     = depth,
            pressure      = -999,
            sound_vel     = -999,  
            depth_msw_r   = depth)
      } else {
        # Else, process CTD data and calculate oxygen saturation
        # Calculate depth from pressure in SEAWATER (factors-in gravity and latitude)
        DAT.tmp <- DAT.tmp %>% 
          mutate(depth      = calc_depth(lat_s, pressure),
                 depth_msw  = depth - altitude,
                 oxygen_sat = calc_sat(salinity, temperature, oxygen_conc))
      }
      
      # Smooth ROV roll, pitch and altitude for width/area calculations
      # Replace NA values with raw (unsmoothed) data
      # Set bad (negative) altitude data to NA
      DAT.tmp$altitude[which(DAT.tmp$altitude < 0)] <- NA
      
      # If altitude data is present (not all values > 0)
      if (length(which(is.na(DAT.tmp$altitude) == F)) > 0) {
        # use linear interpolation to replace NAs
        DAT.tmp$altitude <- as.numeric(na.interp(DAT.tmp$altitude))
        # smooth altitude data
        DAT.tmp$altitude_sm <- as.numeric(ma(DAT.tmp$altitude, order = nav.smoother))
        # replace NA data with non-smoothed data
        isna <- which(is.na(DAT.tmp$altitude_sm) == T)
        DAT.tmp$altitude_sm[isna] <- DAT.tmp$altitude[isna] 
      } else {
        DAT.tmp$altitude_sm <- as.numeric(NA) 
      }
      
      # Smooth pitch data --------------------
      # Set negative (bad) altitude data to NA
      DAT.tmp$pitch[which(DAT.tmp$pitch >= 0)] <- NA
      
      # Use linear interpolation to replace NAs
      DAT.tmp$pitch    <- as.numeric(na.interp(DAT.tmp$pitch))
      DAT.tmp$pitch_sm <- as.numeric(ma(DAT.tmp$pitch,order = nav.smoother))
      
      # Replace NAs with non-smoothed data
      isna <- which(is.na(DAT.tmp$pitch_sm) == T)
      DAT.tmp$pitch_sm[isna] <- DAT.tmp$pitch[isna]
      DAT.tmp$pitch[which(DAT.tmp$pitch > 0)] <- NA
      
      # Smooth roll data ---------------------
      DAT.tmp$roll_sm <- as.numeric(ma(DAT.tmp$roll,order = nav.smoother))
      
      # Replace NAs with non-smoothed data
      isna <- which(is.na(DAT.tmp$roll_sm) == T)
      DAT.tmp$roll_sm[isna] <- DAT.tmp$roll[isna]
      
      # Calculate width and area -------------
      # Uses the method described in Stierhoff et al. (in review) and calc_width function in surveyR package
      width.df <- calc_width(DAT.tmp$pitch_sm, DAT.tmp$altitude_sm, ROV = ROV) 
      
      DAT.tmp <- bind_cols(DAT.tmp, width.df) %>% 
        select(nav_id, object_id, dive_name, date_time, 
               lat_s, long_s, northing_s, easting_s,
               heading_s, cmg_s, speed_s, disp_s,
               lat_r, long_r, depth, northing_r, easting_r,
               heading_r, cmg_r, speed_r, pitch, roll,
               temperature, conductivity, pressure, salinity, 
               sound_vel, oxygen_conc, oxygen_sat,
               altitude, disp_r, good, 
               depth_msw, altitude_sm, pitch_sm, roll_sm)
      
      # Write DAT data frame to CSV file for import into the database
      DAT.write <- DAT.tmp %>% 
        mutate(date_time = format(date_time, format = "%m/%d/%Y %H:%M:%S")) %>% 
        select(-c(depth_msw, altitude_sm, pitch_sm, roll_sm))
      
      write_csv(DAT.write, file.path(i, paste(dive.name, "NavData.txt", sep = "_")), 
                na = "-999")
      
      write_csv(DAT.write, here("Output", paste(dive.name, "NavData.txt", sep = "_")), 
                na = "-999")
      
      # Add results to DAT.temp
      DAT <- bind_rows(DAT, DAT.write)
      
      # Plot CTD data -------------------------------
      ctd.gg <- DAT.tmp %>% 
        select(date_time, depth, depth_msw, salinity, conductivity, temperature, pressure,
               sound_vel, oxygen_conc, oxygen_sat) %>% 
        gather(variable, value, -date_time) 
      
      ctd.p 	<- ggplot(ctd.gg, aes(x = date_time, y = value, group = variable)) + 
        geom_line(colour = "blue") + scale_x_datetime() + 
        facet_wrap(~variable, scales = "free_y") + 
        labs(title = paste("CTD Data from transect ",dive.name,"\n",sep = "")) + 
        xlab("\nTime of Day (GMT)") + ylab("Sensor Value\n") +
        theme_bw() +
        theme(strip.text.x = element_text(size = 12, face = "bold"),
              strip.background = element_rect(fill = "white"),
              panel.spacing = unit(1, "lines"))
      
      # Save CTD plot
      ggsave(ctd.p, filename = here("Figs", paste(dive.name,"PhysData.png", sep = "_")), 
             height = 9.8, width = 13.3, units = "in")
      
      ggsave(ctd.p, filename = file.path(i, paste(dive.name,"PhysData.png", sep = "_")), 
             height = 9.8, width = 13.3, units = "in")
      
      # Plot results from center width and area calculations ----------------------
      # Plot pitch data
      p.gg <- DAT.tmp %>% 
        select(date_time, pitch, pitch_sm) %>% 
        gather(variable, value, -date_time)
      
      p.p	<- ggplot(p.gg, aes(x = date_time, y = value, colour = variable)) + 
        geom_line() +
        labs(title = paste("ROV data from ", dive.name,"\n", sep = "")) +
        xlab("") + ylab("Pitch (degrees)\n") + 
        scale_x_datetime() + 
        scale_colour_manual("Pitch", values = c("black","green"))
      
      # plot roll data
      r.gg <- DAT.tmp %>% 
        select(date_time, roll, roll_sm) %>% 
        gather(variable, value, -date_time)
      
      r.p	<- ggplot(r.gg, aes(x = date_time, y = value, colour = variable)) + 
        geom_line() + 
        xlab("") + ylab("Roll (degrees)\n") + 
        scale_x_datetime() + 
        scale_colour_manual("Roll",values = c("black","blue"))
      
      # plot altitude data
      a.gg <- DAT.tmp %>% 
        select(date_time, altitude, altitude_sm) %>% 
        gather(variable, value, -date_time)
      
      if (length(which(is.na(a.gg$value) == FALSE)) > 0) {
        a.p	<- ggplot(a.gg, aes(x = date_time, y = value, colour = variable)) + 
          geom_line() + 
          xlab("") + ylab("Altitude (meters)\n") + 
          scale_x_datetime() + 
          scale_colour_manual("Altitude", values = c("black","red","blue"))
      }
      
      # plot of ROV depth and altitude data
      z.gg <- DAT.tmp %>% 
        select(date_time, depth, depth_msw) %>% 
        gather(variable, value, -date_time)
      
      z.p <- ggplot(z.gg, aes(x = date_time, y = value, colour = variable)) + 
        geom_line() + 
        scale_colour_manual("Depth", values = c("green", "black")) + 
        xlab("\nTime of Day") + ylab("Depth (m)\n") +	
        scale_x_datetime()
      
      # Save the plot grid
      if (length(which(is.na(a.gg$value) == FALSE)) > 0) {
        nav.grid <- plot_grid(p.p, r.p, a.p, z.p, nrow = 4, align = "v")
        
        save_plot(file.path(i, paste(dive.name, "NavData.png", sep = "_")), 
                  nav.grid, ncol = 1, nrow = 4, base_aspect_ratio = 4)
        
        save_plot(here("Figs", paste(dive.name, "NavData.png", sep = "_")), 
                  nav.grid, ncol = 1, nrow = 4, base_aspect_ratio = 4)
        
      } else {
        nav.grid <- plot_grid(p.p, r.p, z.p, nrow = 3, align = "v")
        
        save_plot(file.path(i, paste(dive.name, "NavData.png", sep = "_")), 
                  nav.grid, ncol = 1, nrow = 3, base_aspect_ratio = 4)
        
        save_plot(here("Figs", paste(dive.name, "NavData.png", sep = "_")), 
                  nav.grid, ncol = 1, nrow = 3, base_aspect_ratio = 4)
      }
      
      # Plot camera data ---------------------------------------------------
      # If camera altitude data is not all NA
      if (length(which(is.na(width.df$camera_altitude) == FALSE)) > 0) {
        width.df$date_time <- DAT.tmp$date_time
        
        w.gg <-  gather(width.df, variable, value, -date_time)
        
        w.p <- ggplot(w.gg, aes(x = date_time, y = value, colour = variable)) + 
          geom_line() + 
          scale_colour_manual("Variable", values = c("green","black","blue")) + 
          xlab("\nDate/time") + ylab("Variable (m)\n") +	
          scale_x_datetime() +	
          labs(title = paste0("Camera data from ", dive.name, "\n"))
        
        ggsave(w.p, filename = file.path(i, paste(dive.name, "CameraData.png", sep = "_")),
               height = 5, width = 10, units = "in")
        
        ggsave(w.p, filename = here("Figs", paste(dive.name, "CameraData.png", sep = "_")),
               height = 5, width = 10, units = "in")
      }
    }
    
    # Process WinFrog events from LOG file -------------------------------------
    log.info <- file.info(file.path(i, 'logs.LOG'))
    
    if (log.info$size == 0) {
      # If the LOG file is empty, do nothing
      cat(paste("Log file for", dive.name, "is empty.\n"))
      
    } else {
      # Else, process the LOG file
      LOG.tmp <- read_csv(file.path(i, "logs.LOG"), col_names = FALSE) %>% 
        select(c(1, 3:11, 46:53))
      
      names(LOG.tmp) <- c("comment","date_time","lat_r","long_r","depth","heading_r","cmg_r","speed_r","northing_r","easting_r",
                          "lat_s","long_s","depth_s","heading_s","cmg_s","speed_s","northing_s","easting_s")
      
      # create nav_id and object_id
      LOG.tmp <- LOG.tmp %>% 
        mutate(event_id  = seq(event.seed, event.seed + n() - 1, 1),
               object_id = seq(1, n(), 1),
               nav_id    = NA_real_,
               lag_s     = NA_real_,
               dive_name = dive.name,
               date_time = mdy_hms(date_time),
               lat_r     = winfrog2dd(lat_r),
               long_r    = winfrog2dd(long_r),
               lat_s     = winfrog2dd(lat_s),
               long_s    = winfrog2dd(long_s))
      
      # Increment event.seed
      event.seed <- max(LOG.tmp$event_id) + 1
      
      if (nrow(DAT.tmp) > 0) {
        # sync log and nav datetimes
        for (kk in 1:nrow(LOG.tmp)) {
          # calculate the time difference between j-th video obs and each nav record
          time.diff	          <- abs(difftime(LOG.tmp$date_time[kk], DAT.tmp$date_time, units = "secs"))
          LOG.tmp$nav_id[kk]	<- DAT.tmp$nav_id[which.min(time.diff)]
          LOG.tmp$lag_s[kk] 	<- min(time.diff)
        }
      }
      
      # select nav data to output to the database
      LOG.write <- LOG.tmp %>% 
        select(event_id, nav_id, lag_s, dive_name, date_time, comment, lat_s, long_s, 
               northing_s, easting_s, heading_s, cmg_s, speed_s,
               lat_r, long_r, depth, northing_r, easting_r, heading_r, cmg_r, speed_r) %>% 
        mutate(date_time = format(date_time, format = "%m/%d/%Y %H:%M:%S"))
      
      # write dive events to text file
      write_csv(LOG.write, file.path(i, paste(dive.name,"DiveEvents.txt", sep = "_")), 
                na = "-999")
      
      write_csv(LOG.write, here("Output", paste(dive.name,"DiveEvents.txt", sep = "_")), 
                na = "-999")
      
      # add results to LOG.temp
      LOG <- bind_rows(LOG, LOG.write)
    }
    
    if (process.photos) {
      # Process photos to get metadata and sync with nav data ####
      photo.list <- dir_ls(here("Data/PHOTOS"), recurse = FALSE, regexp = "\\d{2}-\\d{3}\\w")
      
      p <- photo.list[grep(dive.name, photo.list)]
      
      # If photos were taken, then the photo directory exists in photo.root
      if (length(grep(dive.name, p)) > 0) { 
        # Apply time fix to each photo using ExifTool
        if (.Platform$OS.type == "unix") {
          # System(paste("exiftool -AllDates", timefix, " ", p, sep = ""), intern = T) 
          exiftool_call(paste0("-AllDates", timefix, " ", p))
        } else {
          # System(paste("C:/SOFTWARE/exiftool.exe -AllDates", timefix, " ", p, sep = ""), intern = T)
          exiftool_call(paste0("-AllDates", timefix, " ", p))
        }
        
        # Format photo info
        PHOTO.tmp <- read_exif(p, tags = c("filename","createdate"), recursive = T) %>% 
          rename(filename = FileName) %>% 
          mutate(
            date_time = ymd_hms(CreateDate),
            photo_id  = seq(photo.seed, photo.seed + n() - 1, 1),
            nav_id    = NA_real_,
            lag_s     = NA_real_,
            dive_name = dive.name,
            filepath  = paste("\\\\swc-storage1\\ROV\\PHOTOS", dive_name, filename, sep = "\\"),
            nav_id    = NA,
            lag_s     = NA,
            comment   = NA)
        
        # Increment photo.seed
        photo.seed <- max(PHOTO.tmp$photo_id) + 1
        
        # Sync photos with nav date time
        if (nrow(DAT.tmp) > 0) {
          for (jj in 1:nrow(PHOTO.tmp)) {
            # Calculate the time difference between j-th video obs and each nav record
            time.diff	<- abs(difftime(PHOTO.tmp$date_time[jj], DAT.tmp$date_time, units = "secs"))
            PHOTO.tmp$nav_id[jj]	<- DAT.tmp$nav_id[which.min(time.diff)]
            PHOTO.tmp$lag_s[jj] 	<- min(time.diff)
          }
        } 
        
        # Format for database import
        PHOTO.write <- select(PHOTO.tmp, photo_id, nav_id, lag_s, dive_name, 
                              date_time, filename, filepath, comment) %>% 
          mutate(date_time = format(date_time, format = "%m/%d/%Y %H:%M:%S"))
        
        # Write PHOTO_INFO to a text file
        write_csv(PHOTO.write, file.path(i, paste0(dive.name, "_PhotoInfo.txt")),
                  na = "")
        
        write_csv(PHOTO.write, here("Output", paste0(dive.name, "_PhotoInfo.txt")),
                  na = "")
        
        # Add results to PHOTO.temp
        PHOTO <- bind_rows(PHOTO, PHOTO.write)
        
      } else {
        cat(paste("No photos to process for", dive.name, ".\n"))
        
      }
    }
    
    if (nrow(DAT.tmp) > 0) {
      # Plot lat/lon data of the ROV, ship, and photos -------------------
      # Subset nav data for ship
      ship <- select(DAT.tmp, lat_s, long_s) %>% 
        mutate(vessel = "Sship") %>% 
        rename(lat = lat_s,
               long = long_s)
      
      # Subset nav data for ROV
      rov <- select(DAT.tmp, lat_r, long_r) %>% 
        mutate(vessel = "ROV") %>% 
        rename(lat = lat_r,
               long = long_r)
      
      ll.gg	<- bind_rows(ship, rov)
      
      ll.p	<- ggplot(ll.gg, aes(x = long, y = lat, colour = vessel)) +
        geom_path() + 
        coord_map() + 
        labs(title = paste("Lat/Long data from ", dive.name,"\n", sep = "")) +
        xlab("\nLongitude") + ylab("Latitude\n") + 
        scale_colour_manual("Vessel", values = c("green","black"))
      
      ggsave(ll.p, filename = file.path(i, paste(dive.name, "LatLongData.png", sep = "_")))

      ggsave(ll.p, filename = here("Figs", paste(dive.name, "LatLongData.png", sep = "_")))
    }
    
    # Process CTD cast data --------------------------------------------
    # list CTD files in directory
    CTD.dir <- dir_ls(i, regexp = "\\d{2}-\\d{3}\\w{1}_(CTD.)(\\(\\d{3}-\\d{6}\\))?.DAT")
    
    if (length(CTD.dir) == 0) {
      # If no CTD files are present, do nothing
      cat(paste("No CTD casts to process for ", dive.name, ".\n", sep = ""))
    } else {
      # else process each CTD file
      for (ii in CTD.dir) {
        # Get CTD name
        CTD.name <- str_extract(ii,'\\d{2}-\\d{3}\\w{1}_(CTD.)')
        CTD.tmp <- read_csv(ii, col_names = FALSE)
        CTD.tmp <- CTD.tmp[ , c(1:47,65)]
        # Add variable names to data frame
        names(CTD) <- c("object_id","blank","date_time","lat_c","long_c","depth","northing_c","easting_c",
                        "blank","blank","blank",
                        "heading_r","cmg_r","speed_r","blank","blank","blank","blank","blank","blank","blank","blank",
                        "pitch","roll","temperature","conductivity","pressure","salinity","sound_vel","oxygen_conc",
                        "oxygen_sat","altitude","blank","blank","lat_s","long_s","blank","northing_s","easting_s",
                        "blank","blank","blank","heading_s","cmg_s","speed_s","blank","blank","blank")
        
        # remove variables with no data
        CTD.tmp <- CTD.tmp[ , names(CTD.tmp) != "blank"] %>% 
          mutate(
            ctd_id    = seq(ctd.seed, ctd.seed + n() - 1, 1), 
            object_id = seq(1, n(), 1),
            ctd_name  = str_extract(ii, "\\d{2}-\\d{3}\\w{1}_(CTD.)"),
            dive_name = dive.name,
            date_time = mdy_hms(date_time),
            lat_c     = winfrog2dd(lat_c),
            long_c    = winfrog2dd(long_c),
            lat_s     = winfrog2dd(lat_s),
            long_s    = winfrog2dd(long_s),
            time.int  = c(date_time[2:length(date_time)] - date_time[1:length(date_time) - 1], 0))
        
        # Increment ctd.seed
        ctd.seed <- max(CTD.tmp$ctd_id) + 1
        
        # Process CTD data
        if (ctd.on == FALSE) {  
          # If CTD not present, make data -999
          CTD.tmp <- CTD.tmp %>% 
            mutate(
              sal           = -999,
              conductivity  = -999,
              oxygen_conc   = -999,
              oxygen_sat    = -999,
              depth_p_c     = depth,
              pressure      = -999,
              sound_vel     = -999,  
              depth_msw_c   = depth)
          
        } else {
          # Else, process CTD data and calculate oxygen saturation
          # Calculate depth from pressure in SEAWATER (factors-in gravity and latitude)
          CTD.tmp <- CTD.tmp %>% 
            mutate(
              depth         = calc_depth(lat_s, pressure),
              depth_msw	    = depth - altitude,
              oxygen_sat    = calc_sat(salinity, temperature, oxygen_conc))
        }
        
        # select CTD data to output to the database
        CTD.output <- CTD.tmp %>% 
          select(ctd_id, object_id, dive_name, ctd_name, date_time, lat_c, long_c, 
                 depth, temperature, conductivity, pressure, salinity, sound_vel,
                 oxygen_conc, oxygen_sat) %>% 
          mutate(date_time = format(date_time, format = "%m/%d/%Y %H:%M:%S"))
        
        # Save CTD results to file					
        write_csv(CTD.output, file.path(i, paste0(CTD.name,".txt")), na = "-999")
        
        write_csv(CTD.output, here("Output", paste0(CTD.name,".txt")), na = "-999")
        
        # add CTD results to CTD.temp
        CTD <- bind_rows(CTD, CTD.output)
        
        # plot data from each CTD cast
        CTD.gg <- select(CTD.tmp, object_id, temperature, salinity, oxygen_conc, sound_vel, depth) %>% 
          rename(
            "Temperature (C)"             = temperature,
            "Salinity (PSU)"              = salinity,
            "Oxygen Concentration (uMol)" = oxygen_conc,
            "Sound Velocity (m/s)"        = sound_vel,
            Depth                         = depth) %>% 
          gather(variable, value, -object_id, -Depth)
        
        ctd.p <- ggplot(CTD.gg, aes(value, Depth, colour = variable)) +
          geom_path() + 
          facet_wrap(~variable, nrow = 1, scales = "free_x") + 
          scale_y_continuous("Depth (m)\n", lim = c(min(CTD.gg$Depth), 0), expand = c(0,0)) + 
          xlab("\nSensor Value") + 
          theme(legend.position  = "none",
                strip.text.x     = element_text(size = 12, face = "bold"),
                strip.background = element_rect(fill = "white"),
                panel.spacing    = unit(1, "lines")) + 
          ggtitle(paste("CTD profile for", CTD.name, "\n"))
        
        ggsave(ctd.p, filename = file.path(i, paste0(CTD.name, ".png")),
               height = 9.8, width = 13.3, units = "in")
        
        ggsave(ctd.p, filename = here("Figs", paste0(CTD.name, ".png")),
               height = 9.8, width = 13.3, units = "in")
      }
    }
  }
  
  # Create a vector of processed directories
  processed.dat <- unique(file.path(dat.root, DAT$dive_name))
  saveRDS(processed.dat, here("Output/processed_dat.rds"))
}
```

# Plot results

```{r plot-results}
# Create shapefiles of dive tracks
transects <- DAT %>% 
  st_as_sf(coords = c("long_r","lat_r"), crs = 4326) %>% 
  group_by(dive_name) %>% 
  summarise(do_union = F) %>% 
  st_cast("LINESTRING") 

st_write(transects, here("Output", paste(dir.start, dir.end, "Transects.shp", sep = "_")), 
         delete_layer = TRUE)
```


# Export results

```{r export-results}
# Create output directories
dir_create(here("Output/_Results"), c("CameraData","CTD_Casts","LatLongData","PhysData","NavData"))

# Copy plots to Results directory
# Camera data
file_copy(dir_ls(here("Data/NAVDATA"), regexp = "*CameraData.png", recurse = TRUE),
          here("Output/_Results/CameraData"),
          overwrite = TRUE)
# CTD data
file_copy(dir_ls(here("Data/NAVDATA"), regexp = "*CTD.png", recurse = TRUE),
          here("Output/_Results/CTD_Casts"),
          overwrite = TRUE)
# LatLong data
file_copy(dir_ls(here("Data/NAVDATA"), regexp = "*LatLongData.png", recurse = TRUE),
          here("Output/_Results/LatLongData"),
          overwrite = TRUE)
# Nav data
file_copy(dir_ls(here("Data/NAVDATA"), regexp = "*NavData.png", recurse = TRUE),
          here("Output/_Results/NavData"),
          overwrite = TRUE)
# Physical data
file_copy(dir_ls(here("Data/NAVDATA"), regexp = "*PhysData.png", recurse = TRUE),
          here("Output/_Results/PhysData"),
          overwrite = TRUE)

# Write NAV data for all dives to text file
write_csv(DAT, file = here("Output/_Results", paste(dir.start, dir.end,"NavData.txt", sep = "_")),
          na = "-999")

# Write PHOTO data for all dives to text file
if (nrow(PHOTO) > 0) {
  write_csv(PHOTO, file = here("Output/_Results", paste(dir.start, dir.end,"PhotoInfo.txt", sep = "_")),
            na = "")
}

# Write LOG data for all dives to text file
if (nrow(LOG) > 0) {
  write_csv(LOG, file = here("Output/_Results", paste(dir.start, dir.end,"DiveEvents.txt", sep = "_")),
            na = "-999")
}

# Write CTD data for all dives to text file
if (nrow(CTD) > 0) {
  write_csv(CTD, file = here("Output/_Results", paste(dir.start, dir.end,"CTDCasts.txt", sep = "_")),
            na = "-999")
}
```

# Report new seeds

```{r eval=FALSE}
# Calculate and report new indices for next transect ####
# Calculate next nav_id
nav.seed <- max(DAT$nav_id) + 1	
# Calculate next event_id
if (log.info$size > 0) {
  # if log file is not empty, increment the event seed by 1
  event.seed <- max(LOG$event_id) + 1
} else {
  # else keep the same
  event.seed <- event.seed
}
# Calculate next photo_id
if (length(grep(dive.name, p)) > 0) {
  # if the dive name is in the list of photo directories, increment the photo seed by 1
  photo.seed <- max(PHOTO$photo_id) + 1
} else {
  # else keep the same
  photo.seed <- photo.seed
}
# Calculate next ctd_id
if (length(CTD.dir) > 0) {
  # if CTD casts were performed, increment the CTD seed by 1
  ctd.seed <- max(CTD$ctd_id) + 1
} else {
  # else keep the same
  ctd.seed <- ctd.seed
}

# Report new indices for next dive ####
cat(paste("Finished processing dives", dir.start, "to", dir.end, "\n", sep = " "),
    paste("Next nav_id is  : ", nav.seed, "\n"),
    paste("Next event_id is: ", event.seed, "\n"),
    paste("Next photo_id is: ", photo.seed, "\n"),
    paste("Next ctd_id is  : ", ctd.seed,"\n"), sep = "")
```

